```{r}
renv::activate()
source("./packages.R")
```

We will merge all the data from 2022. 
For that, first we will check the directories

```{r}
library(hms)
dirs <- dir_ls(here(c("data/datos2022")))
dirs
```

Now we are going to get the list of files of each dir. 
```{r}
files <- dirs %>% map(dir_ls, glob = "*collar*")
files

```


Now we are going to read and merge all these files into a single tibble for labelling
Each folder contains data from a single sheep, which is splitted into various data files

We have to make a proper datetime variable for an inter sheep merge. 

```{r}
library(chron)
files1 <- as.list(files[[1]])
files1


# data <- list()
# for (sheep in seq_along(files)) {
#    data[[sheep]] <-  as.list(files[[sheep]]) %>% map(.f = ~{fread(.x ) %>% as_tibble}) %>% bind_rows()
# }


vars <- c("Pitch angle", "Time_dd:hh:mm:ss.sss", "Decimal secs", "Date")

data <-  as.list(files1) %>% map(.f = ~{fread(.x, select=vars) %>% as_tibble}) %>% bind_rows() %>% 
    rename(hms_time=`Time_dd:hh:mm:ss.sss`) %>% mutate(Date = lubridate::dmy(Date))  %>% 
    mutate(hms_time=as.POSIXct(data$hms_time, format="%H:%M:%S"))  %>% 
    mutate(hour=lubridate::hour(hms_time)) %>% 
    mutate(minute=lubridate::minute(hms_time)) %>% 
    mutate(second=lubridate::second(hms_time)) %>% 
    select(-hms_time) %>% 
    mutate(year=lubridate::year(Date)) %>% 
    mutate(month=lubridate::month(Date)) %>% 
    mutate(day=lubridate::day(Date))


data

data2 <- data  %>% 
    mutate(hms_time=as.POSIXct(data$hms_time, format="%H:%M:%S")) 
    
    
data3 <- data2 %>% 
    mutate(hour=lubridate::hour(hms_time)) %>% 
    mutate(minute=lubridate::minute(hms_time)) %>% 
    mutate(second=lubridate::second(hms_time)) %>% 
    select(-hms_time) 
    
data4 <- data3 %>% 
    mutate(year=lubridate::year(Date)) %>% 
    mutate(month=lubridate::month(Date)) %>% 
    mutate(day=lubridate::day(Date))


data4


data_head <- head(data)
data_head

data %>% glimpse

data$hms <- chron(times=data$hms)

lubridate::hour(data_head$hms)



data_head

test <- as.POSIXct(data_head$hms,format="%H:%M:%S")

lubridate::hour(test)


hms %>% str



data  %>% glimpse
ymd(data$Date)

hms <- data_head$hms

hms



lubridate::hour(hms)
test <- as.POSIXct(hms,origin="1970-01-01")
test


library(chron)
x <- chron(times=hms)
x

data_head <- head(data)
data_head <- data_head %>% mutate(Date = dmy(Date))
data_head
lubridate::year(data_head$Date)
make_datetime(
    lubridate::year(data_head$Date), 
    lubridate::month(data_head$Date), 
    lubridate::day(data_head$Date), 
    lubridate::hour(data_head$hms), 
    lubridate::minute(data_head$hms), 
    lubridate::hour(data_head$hms)
)


lubridate::day(data_head$Date)
hms::hms(as.character(as.numeric((data_head$`Time_dd:hh:mm:ss.sss`)))


data_head_date <- data_head$Date
dmy(data_head_date)

colnames(data[[1]])

bind_cols(data)
test <- map2(data[[1]], data[[2]], left_join)


## for merging  https://stackoverflow.com/questions/5965698/merge-unequal-dataframes-and-replace-missing-rows-with-0




```




```{r}

vars <- c("Pitch angle", "Time_dd:hh:mm:ss.sss", "Decimal secs", "Date")
vars_1 <- c("Pitch angle")
data_raw <- list()
data <- list()

for (i in seq_along(dirs)){
    for (j in seq_along(files[[i]])){
        data_raw[i][[j]] <-  files[[i]][[j]] %>% fread(select = vars) %>% as.data.frame  %>% as_tibble
        
    }
    #`data[[i]] <- bind_rows()
}


data1 <- files[[1]][[2]] %>% fread(select = vars) %>% as.data.frame  %>% as_tibble


map(files[[1]]as.as.list())

# this is for when merging inter sheep files, not intra sheep, for intrasheep we need bindrows
data_raw_t <- files[[1]][[1]] %>% fread(select = vars) %>% as.data.frame  %>% as_tibble %>% dplyr::rename('Pitch_angle_1' = 'Pitch angle')
data_raw_t1 <- files[[1]][[2]] %>% fread(select = vars_1) %>% as.data.frame  %>% as_tibble %>% dplyr::rename('Pitch_angle_2' = 'Pitch angle')


.x  <- data_raw[1][[1]]
.y  <- data_raw[1][[2]] %>% select('Pitch angle')
temp = bind_columns(.x, .y)
for (j in seq_along(files[[1]])) {
    temp = bind_cols(data_raw[[1]][[j]], data_raw[[1]][[j+1])
    data = 
}


.x  <- data_raw[1][[1]]
.y  <- data_raw[1][[1]]

str(data_raw[1])
test3 <- dplyr::bind_rows(data_raw[1], .id = 'Pitch angle')



glimpse(test3)

test <- bind_cols(.x, .y)

test2 <- reduce(data_raw[1], bind_cols)

test4 <- bind_cols(data_raw[[1]])

test4  %>% glimpse

data_raw[[1]][[1]] %>% glimpse

test2 %>% glimpse

test %>% glimpse


qplot(data_raw[[1]][[1]])



```
 ```{r}
## Read text file with readr


.x <- files[[1]][[1]]
.x

vars <- c("Pitch angle", "Time_dd:hh:mm:ss.sss", "Decimal secs")


data_raw <- list()
for (i in seq_along(dirs)){
    for (j in seq_along(files[[i]])){
        data_raw[i][[j]] <-  files[[i]][[j]] %>% fread(select = vars) %>% as.data.frame 
        print(j)
    }
}




data_raw %>% str


for (i in seq_along(files)){
    glimpse(data_raw[[1]][[i]])
}

glimpse(data_raw)

data_raw[[1]] %>% glimpse


map(files[[1]], .f = ~ {fread(.x, select = vars) %>% as.data.frame}



test <- read.table(.x, sep = "\t", header = TRUE)

test <- read_tsv(.x, sep = "\t", header = TRUE)

map(files[[1]], ~ read.table(.x, sep = "\t", header = TRUE))







data_raw <- list()
for (dir in seq_along(dirs)){
    map(files[[dir]], ~ read.table(.x, sep = "\t", header = TRUE))
}


data_raw <- list()
for (dir in seq_along(dirs)) {
    for (file in seq_along(files)) {
        data_raw[dir][[file]] <- files[[dir]][[file]] %>% read.table(sep = "\t", header = TRUE)       
    }
} 


data_raw %>% glimpse



.x

```